% Acknowledgements (Acceptable American spelling) are optional
% So are Acknowledgments (proper English spelling)
    \chapter*{Acknowledgments}
	\epigraph{Some people are of the opinion that acknowledgments ought to be concise, relevant to the thesis, and devoid of sappy sentiment. They are probably right, but they aren't the ones writing this.}{Varchas Gopalaswamy}

This thesis wouldn't have existed without my thesis advisor, Daniel Borrero. Thank you, Daniel, for suggesting an awesome thesis topic, for motivating me, for spending a ton of time correcting more terrible thesis drafts than any human should have to, for introducing me to dynamical systems theory, and for all the grad school help. Good luck with setting up the Taylor-Couette system.\\

I would also like to thank John Gibson from the University of New Hampshire for making this thesis even remotely feasible by writing {\tt Channelflow}, and for taking the time to correspond with me via email and Hangout. Your advice has been invaluable. \\
 
By definition, this thesis wouldn't have existed were I not a physics major, so I'd like to thank the department as a whole for being the greatest department at Reed. Thank you, Lucas, for being an extremely supportive academic advisor and for a  challenging junior year. Thank you, Joel, for introducing me to the world of scientific computation, for running physics softball, and for your help with this thesis. Thank you, Darrell, for reminding me why quantum mechanics is awesome. Thank you, Johnny, for all the help with the grad school process. Thanks also to the physics seniors -- shoutout to Julia and Neal for their tenure as the Pub Czars, Taras for his tenure as the Cookie Czar, Dan for putting up with the frantic late night calls for help, and Newton for being a rad office buddy.  I will be proud to say that I was once a part of this group. You guys have made my time here special. \\

To Amma and Appa -- Thank you for all your love and support throughout all these years, and always being understanding and there for me when I have the occasional mental breakdown. Also, thanks for going ``Yes, let's send Varchas to this college we've literally never heard anything about, this sounds like a great idea." I hope you feel that you've made the right decision. 

Dear Medha: Thank you for allowing me to lecture at you about stuff you may or may not actually be interested in, for enabling me in some truly horrendous jokes, for doing all the housework whenever I would come back for vacation, and generally being an awesome sister. \\

To all my friends -- Thank you for making Reed the best of times, and for supporting me through the worst of times. Words cannot express how much I will miss you all.\footnote{Yes Matt, even you.}




